---
title: "US Tilapia Imports"
author: "Erica Dale"
date: "2022-08-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


For Task 2, you will work with the us_tilapia_imports.csv. It exists in your eds221-day5-comp project - copy that file into your data folder for your day 5 activities project. The data are from the USDA Economic Research Service, and contain annual total volumes (in thousands of pounds) of tilapia imports to the United States from different countries.

You can decide if you want to do this all in separate steps, or piped together in sequence, or some combination. Make sure if you pipe things together, you check the output at every step.

Create a new .Rmd in your docs folder called us_tilapia_imports.Rmd

Attach the tidyverse, here and janitor packages

Read in the data as us_tilapia_imports

```{r}
library(tidyverse)
library(here)
library(janitor)
us_tilapia_imports <- read_csv(here("data", "us_tilapia_imports.csv"))
summary(us_tilapia_imports)
names(us_tilapia_imports)
dim(us_tilapia_imports)
```

Explore the data. What are the classes of the columns? Remember some tools we’ve used: summary, names, dim, skim, etc.

Use pivot_longer() to reshape the data into long format (currently, the variable “year” is spread across multiple columns). Remember to store the output so you will be able to use the reshaped data this creates.

Check the class of the year column. What is it, and why do you think that’s the case? Then, coerce the year column to numeric (e.g. using mutate() and as.numeric() in combination)

Use dplyr::group_by() %>% summarize() to find the total US tilapia imports by year, store as yearly_tilapia_tot

```{r}
us_tilapia_clean <- pivot_longer(data = us_tilapia_imports, cols = -country, values_to = "TilapiaCount", names_to = "Year") |> 
  mutate(NewYear = as.numeric(us_tilapia_clean$Year))

yearly_tilapia_tot <- us_tilapia_clean |> 
  group_by(NewYear) |> 
  summarize(yearly_import = sum(TilapiaCount, na.rm = TRUE))

yearly_tilapia_tot
```

Create a ggplot line graph of total US tilapia imports for all years in yearly_tilapia_tot. Update axis labels (include units as necessary), then export your graph as a .png to figs.

```{r}
ggplot(data = yearly_tilapia_tot, aes(x = NewYear, y = yearly_import)) +
  geom_line() +
  labs(x = "Years", y = "Total import")
```


Create a subset that only retains imports from Ecuador, Honduras, Costa Rica, and Mexico (you decide what to name this)

```{r}
central_am <- us_tilapia_clean %>%
  filter(country %in% c("Ecuador", "Honduras", "Costa Rica", "Mexico"))
```


Create a ggplot graph of total US tilapia imports over time, for those four countries in the subset you created above, separated by country. Update axis labels, add a title, customize your color scheme, update the theme. Export a .jpg of your graph to figs.

```{r}
ggplot(data = central_am, aes(x = NewYear, y = TilapiaCount)) +
  geom_line(aes(color = country)) +
  facet_grid(vars(country)) +
  labs(title = "Central American Imports", x = "By Year", y = "Fish Count") +
  theme_dark()

ggsave(path = here("figs"), filename = "centralam_country_count.jpg")
```

